Article figures

### 2021-06-11 src/article_figures.Rmd
all_results now includes TPR and FPR as well

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
library(knitr)
library(tidyverse)
save_figs = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

@@

@@

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
                    `Previous Hb deferral` = "Previous donation deferred",
                    `Polygenic risk score` = "Polygenic score",
                    `RNF43 mutant` = "RNF43")
```

@@

## Alternative effect sizes and importances of variables

```{r, include=FALSE}
alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {

@@

```{r}
res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180)
  ggsave(filename="presentation-effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180, height=90)
g
```

@@

all_ids <- c(ids, dummy2_ids)

# get_threshold_df <- function() {
#   
#   t6 <- tibble(
#       Id = factor(all_ids),
#       type = factor("Threshold"),
#       value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
#   t12 <- tibble(
#       Id = factor(all_ids),
#       type = factor("Threshold"),
#       value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
#   )
#   t <- bind_rows(t6, t12)
#   t <- t %>%
#     mutate(low=value, 
#            high=value,
#            sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
#            #model=factor(str_split_fixed(Id, "-", 3)[,3])
#            model=rep(c("t6", "t12"), each=length(all_ids))
#            )
#   return(t)
# }

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  # Note that recode stupidly uses order "old"="new"
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  # s <- map(s, function(x) when(x, 
  #                              x == "progesa" ~ "eProgesa",
  #                              x == "finngen" ~ "Biobank",
  #                              x == "findonor" ~ "FinDonor",
  #                              ~ "koe"))
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
  if (FALSE) {
    df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
    df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
    df <- bind_rows(df1, df2)
    #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
      mutate(type = fct_recode(type, !!!new_names))

    thresholds <- get_threshold_df()
    df <- bind_rows(df, thresholds)
    df <- df %>% 
      mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
  } else {
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")

    df <- df %>% 
      pivot_wider(names_from=c("type"), values_from=value) %>%
      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
             high = ifelse(is.na(high), value, high))

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=case_when(variable == "threshold6" ~ "t6",
                             variable == "threshold12" ~ "t12",
                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")),
             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
             variable = fct_recode(variable, !!!new_names))

@@

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

@@

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs)
  ggsave(filename="performance-forest-plot.pdf", #  title="Performance forest plot", 
         path=fig_path, plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
g
```

@@

```{r}
df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
df <- df %>% 
  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
names(df) <- str_remove(names(df), " value$")
df2 <- df %>%
  select("Id", "AUROC", "AUROC low", "AUROC high", "AUPR", "AUPR low", "AUPR high", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
kable(df2, digits=2)
```

Create a table of atot and TPR and FPR values for each method/data. Q is the fraction of avoided deferrals.

```{r}
#data_frames <- map(dummy2_ids, get_data_frame)
df2 <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
```

```{r}
helper <- function(Id, threshold6, threshold12) {
  df2 <- get_data_frame(Id)
  r6 <- get_rates(df2, threshold6)
  r12 <- get_rates(df2, threshold12)
  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, p = parameters)
  #df <- tibble(a=1, b=2)
  df$Id <- Id
  return(df %>% select(Id, everything()))  
}

d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
result <- pmap_dfr(df2, helper) 
result <- result %>%
  mutate(across(c(q6, q12), function(x) x / d)) %>%
  rename(Q6=q6, Q12=q12)
kable(result, digits=3)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
library(knitr)
library(tidyverse)
save_figs = FALSE
save_tables = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
table_path="~/FRCBS/results-for-eba-paper/table"
all_results_filename <- sprintf("%s/raw-data-2021-06-11.tsv", table_path)
compute=FALSE
```

Get all variables and their confidence intervals

```{r}
if (compute) {
  all_results <- process_all_data(all_ids)
  write_tsv(all_results, file=all_results_filename)
} else {
  all_results <- read_tsv(file=all_results_filename)  
}
```

Same results in wide format xlsx

```{r}
if (save_tables) {
  library(openxlsx)
  temp <- all_results %>% pivot_wider(names_from=c(variable, type))
  filename <- sprintf("%s/pretty-data-2021-06-11.xlsx", table_path)
  openxlsx::write.xlsx(temp, file = filename, overwrite=TRUE)
}
```

```{r Cost surface}
g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"))
if (save_figs) {
  filename <- sprintf("%s/cme_cost_surface_werr.pdf", fig_path)
  ggsave(filename=filename,  width = 180,  height = 120, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
}
g
```

```{r Probability thresholds}
threshold_range <- seq(0.02, 0.98, 0.02)
data_frames <- map(all_ids, get_data_frame)
#get_thresholds <- function(df,  p=parameters, thresholds = seq(0.1, .9, .1), id=id)
thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_thresholds(df, parameters, threshold_range, id))

thresholds2 <- thresholds %>% 
  filter(Id=="progesa-both-rf") %>%
  select(probability, E6, E12) %>%
  pivot_longer(cols = c(E6, E12), names_to = "variable", values_to = "E") %>%
  extract(col="variable", regex = "([a-zA-Z]+)([0-9]+)", into=c("variable", "month")) %>%
  select(-variable)

min_thresholds <- thresholds2 %>%
  group_by(month) %>%
  slice_min(order_by=E, n=1, with_ties = FALSE) %>%
  ungroup()
```

```{r Probability cutoff curve}
g <- thresholds2 %>%
  mutate(month=factor(month, levels=c(6, 12))) %>%
  ggplot(aes(x=probability, y=E, color=month)) +
  geom_line() +
  geom_point(data=min_thresholds, mapping=aes(x=probability, y=E), 
             color="black", size=2, inherit.aes = FALSE) +
  geom_label(data=min_thresholds, mapping=aes(x=probability, y=E, label=sprintf("p >= %.2f : %.2f €", probability, E)), 
             inherit.aes = FALSE, nudge_y=0.3, size=2) +
  ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)") +
  theme(legend.position="bottom")
g
if (save_figs) {
  filename <- sprintf("%s/cme_prob_cut_curve.pdf", fig_path)
  ggsave(filename=filename,  width = 90,  height = 80, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
}
```

@@

This is the OLD visualisation.

@@

```{r Read FinnGen DLMM posteriors, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
                    `Previous Hb deferral` = "Previous donation deferred",
                    `Polygenic risk score` = "Polygenic score",
                    `RNF43 mutant` = "RNF43")
```

@@

## New effect sizes and importances of variables

```{r, include=FALSE}
new_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11) 
{

@@

```{r}
res <- new_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180)
  ggsave(filename="presentation-effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180, height=90)
g
```

@@

#all_ids <- c(ids, dummy_ids)

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  # Note that recode stupidly uses order "old"="new"
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
  if (FALSE) {
    # df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
    # df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
    # df <- bind_rows(df1, df2)
    # #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
    # 
    # df <- df %>%
    #   mutate(Id=factor(Id, levels=rev(all_ids)),
    #          type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
    #          sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
    #          model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
    #   mutate(type = fct_recode(type, !!!new_names))
    # 
    # thresholds <- get_threshold_df()
    # df <- bind_rows(df, thresholds)
    # df <- df %>% 
    #   mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
  } else {
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
    df <- read_tsv(all_results_filename)

    df <- df %>% 
      filter(variable %in% c("AUROC", "AUPR", "F1", "E6", "E12", "threshold6", "threshold12")) %>%
      pivot_wider(names_from=c("type"), values_from=value) %>%
      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
             high = ifelse(is.na(high), value, high))

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=case_when(variable == "threshold6" ~ "t6",
                             variable == "threshold12" ~ "t12",
                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "baseline", "t6", "t12")),
             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
             variable = fct_recode(variable, !!!new_names))

@@

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "baseline"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

@@

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs) {
  filename <- sprintf("%s/performance-forest-plot.pdf", fig_path)
  ggsave(filename=filename, #  title="Performance forest plot", 
         plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
}
g
```

@@

```{r Performances as LaTeX table}
#df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
df <- read_tsv(all_results_filename)
df <- df %>% 
  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
names(df) <- str_remove(names(df), " value$")
df2 <- df %>%
  select("Id", "AUROC", "AUROC low", "AUROC high", "AUPR", "AUPR low", "AUPR high", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
kable(df2, digits=2)
if (save_tables) {
  filename <- sprintf("%s/performance.tex", table_path)
  latex_code <- kable(df2, format="latex", digits=2, 
                      caption="The numeric values of the classification performance metrics and their 95\\% confidence intervals in Fig. 2.", 
                      label="tab:performance",
                      linesep="")
  cat(latex_code, file=filename, sep="\n")
}
```

## Everything below this is unnecessary

Create a table of atot and TPR and FPR values for each method/data. q is the fraction of avoided deferrals.

```{r Get optimal thresholds, eval=FALSE}
#data_frames <- map(dummy2_ids, get_data_frame)
#compute=TRUE
if (FALSE) {
  threshold_range <- seq(0.02, 0.98, 0.02)
  data_frames <- map(all_ids, get_data_frame)
  thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_optimal_thresholds(df, parameters, threshold_range, id)) %>%
    mutate(Id=all_ids) %>%
    select(Id, threshold6, threshold12)
} else {
  optimal_thresholds <- all_results %>% 
    filter(variable %in% c("threshold6", "threshold12")) %>%
    pivot_wider(names_from=variable, values_from = value) %>%
    select(-type)
}
```

```{r, eval=FALSE}
helper <- function(Id, threshold6, threshold12) {
  df3 <- get_data_frame(Id)
  r6 <- get_rates(df3, threshold6)
  r12 <- get_rates(df3, threshold12)
  sex <- get_sex(Id)
  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, sex=sex, p = parameters)
  #df <- tibble(a=1, b=2)
  df$Id <- Id
  return(df %>% select(Id, everything()))  
}

#d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
result <- pmap_dfr(optimal_thresholds, helper) 
# result <- result %>%
#   mutate(across(c(q6, q12), function(x) x ) %>%
#   inner_join(thresholds, by="Id")
#   #rename(Q6=q6, Q12=q12)
kable(result, digits=3)
```
